---
title: "Automatic Differentiation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Automatic differentiation works by code-generating C++ that uses the CppAD library.

Here is an scalar example:

```{r eval=TRUE}
library(nCompiler)
nc1 <- nClass(
  Cpublic = list(
    Cv = 'numericScalar',
    Cfoo = nFunction(
      fun = function(x) {
        return(1.5*x+1)
      },
      argTypes = list(x = 'numericScalar'),
      returnType = 'numericScalar')
  ),
  enableDerivs = 'Cfoo'
)
set_nOption('automaticDerivatives', TRUE)
nc1_generator <- nCompile_nClass(nc1, interface = "generic")
nc1_obj <- nc1_generator()
method(nc1_obj, "Cfoo")(1.32)## 1.5*1.32 + 1 = 2.98
derivs <- method(nc1_obj, "Cfoo_derivs_")(1.32, c(0, 1)) # 0-th order and 1st-order derivatives
value(derivs, "gradient") # gradient is 1.5
```

Here is a vector example:

```{r}
nc2 <- nClass(
  Cpublic = list(
    Cv = 'numericScalar',
    Cfoo = nFunction(
      fun = function(x) {
        ans <- 1.5 *x
        return(ans)
      },
      argTypes = list(x = 'numericVector(length = 2)'),
      returnType = 'numericVector(length = 2)')
  ),
  enableDerivs = 'Cfoo'
)
set_nOption('automaticDerivatives', TRUE)
nc2_generator <- nCompile_nClass(nc2, interface = "generic")
nc2_obj <- nc2_generator()
method(nc2_obj, "Cfoo")(c(1.48, 1.52)) ## 1.5*c(1.48, 1.52)
derivs <- method(nc2_obj, "Cfoo_derivs_")(c(1.48, 1.52), c(0, 1))
value(derivs, "gradient") ## The gradient is 1.5, 1.5
```

# Future plans

- Use CppAD features to obtain 2nd-order (similar to `nimble`) and higher-order derivatives.
- Test derivatives as we port more math functions from `nimble` to `nCompiler`.  Currently what it can do is limited.
- Ensure that derivatives work across multiple functions.
- Support derivatives for `nFunction` as well as for `nClass` methods.