---
title: "Algorithmic (or Automatic) Differentiation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Automatic differentiation is implemented in `nCompiler` by code-generating C++ that uses the `CppAD` library.

Here is an example, using the simple function `a * exp(v)`.  Note that type declarations need to include fixed lengths.

The system currently needs to be turned on via an option.

```{r}
library(nCompiler)
set_nOption('automaticDerivatives', TRUE)
a_exp_v <- nClass(
  classname = "a_exp_v",
  Rpublic = list(),
  Cpublic = list(
    go = nFunction(
      fun = function(a = 'numericScalar',
                     v = 'numericVector(length = 3)') {
        return(a*exp(v))
      },
      returnType = 'numericVector(length = 3)')
  ),
  enableDerivs = 'go',
)

C_a_exp_v <- nCompile(a_exp_v)
my_C_a_exp_v <- C_a_exp_v$new()
## Here is the Jacobian
nDerivs(my_C_a_exp_v$go(2, 1:3))$gradient
```

The result shows the Jacobian, the matrix of derivatives of each output (column) with respect to each input (row).  The inputs are given in a concatenated order, in this case `a`, `x[1]`, `x[2]`, `x[3]`.

# Future plans

- Support control over which arguments one wants derivatives with respect to.
- Use CppAD features to obtain 2nd-order (similar to `nimble`) and higher-order derivatives.
- Test derivatives as we port more math functions from `nimble` to `nCompiler`.
- Ensure that derivatives work across multiple functions.
- Support derivatives for `nFunction` as well as for `nClass` methods.