---
title: "Extensibility and developer tools"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Developer tools

Developer tools for nCompiler include the following:

- A clearly defined set of compiler stages for turning an `nClass` or `nFunction` to generated C++.
- Ability to control the starting and stopping stage of compilation.
- Ability to generate a log of output of each step taken in compilation.  This is helpful because often an error is generated downstream of where the actual mistake occurs.
- A custom error handler that provides information on the line of source code involved in the error and the specific step of processing where it occurred, with more information than what R would otherwise report.

These tools currently provide different degrees of coverage and control for different compiler stages.

At some point we will add examples of these tools to this page.

# Extensibility

Each compiler stage recurses through the code tree and can call a *handler* function for each function in the code.  Each handler for each stage has a clearly defined task to accomplish.  The set of handlers (and possibly associated control parameters) are looked up from a master list, which has potential to be extended by a user who wants to add support for a new function to be compiled to C++.

An example of the entry defining how "+" will be handled is the following:
```{r}
library(nCompiler)
nCompiler:::operatorDefEnv[['+']]
```

Of course, extending what nCompiler supports will necessitate more knowledge of its internals.

## Future steps

- Support for debugging-mode messages in generated C++.
- Support for a `browser()`-like function from generated C++ to copy local variables into R objects and call R for user inspection.
- Support for providing new compilation handlers.