---
title: "nFunctions and nClasses"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

nCompiler provides new kinds of functions and classes that can be fully or partially compiled.  They can also be used uncompiled (i.e. in R), which is useful for debugging.

# `nFunction`

An `nFunction` is created from an R function and information about argument types and the return type.

```{r load-package}
library(nCompiler)
```

```{r mult2}
mult2 <- function(x) {
  ans <- 2 * x
  return(ans)
}

nf_mult2 <- nFunction(
  fun = mult2,
  argTypes = list(x = 'numericVector'),
  returnType = 'numericVector'
)
nf_mult2(1:3)
```

This can be compiled by:
```{r compile-mult2}
# With a new installation, one might first see ouput from building a local library called nCompLocal.
Cnf_mult2 <- nCompile(nf_mult2)
Cnf_mult2(1:3) # The compiled version works.
```

The vector component-wise multiplication is accomplished using the Eigen library in the code-generated C++.

An `nFunction` can also be created by writing type information in the format of default arguments and providing an explicit `returnType` statement anywhere in the function body.  For example:

```{r mult2-alt}
nf_mult2 <- nFunction(
  fun = function(x = numericVector()) {
    returnType(numericVector())
    ans <- 2 * x
    return(ans)
  }
)
```

# `nClass`

`nClass`es are similar to reference classes and R6 classes, but they can have two categories of fields and methods: those to be handled via C++ compilation and those that will remain pure R.  A `nClass` is implemented as an R6 class, so its syntax follows R6's syntax.

```{r add_class}
add_class <- nClass(
  classname = "add_class",
  Rpublic = list(               # only available in R
     Rnumber = NULL,
     add_Rnumber = function(v) {
       v + Rnumber
     }
     # initialize method not supported through compilation yet.
  ),
  Cpublic = list(                # available in R or compiled version
    # field (aka member data) -- must have a type declaration
    Cnumber = 'numericScalar', 
    # method (aka member function) -- written as an nFunction
    add_Cnumber = nFunction(
      fun = function(v = numericVector()) {
        ans <- v + Cnumber
        return(ans)
      },
      returnType = 'numericVector')
  )
)
```

This can be used uncompiled for debugging purposes.  Then it can be compiled.

For objects of the compiled class, there are two kinds of interfaces, generic and full.  In a generic interface, there isn't an R6 object for every compiled object.  This saves a lot of R processing and memory in large collections of compiled objects.  In a full interface, there is an R6 object for every compiled object.  The full interface is better for nice use of a small number of objects.

Let's compile `add_class`:
```{r compile-add-class}
# Something about the non-standard evaluation of knitr seems
# to cause a problem with nCompile in this context.  The slightly
# lower-level nCompile_nClass works.
Cadd_class <- nCompile_nClass(add_class)
# By default we get a full interface, which
# behaves as one would expect for a class generator.
obj <- Cadd_class$new()
obj$Cnumber <- 3  ## This automatically stores into the C++ object
obj$Cnumber
obj$add_Cnumber(2:4) ## and automatically calls the C++ method.
```

```{r compile-add-class-generic}
# We can also make a generic interface (or get it out of the full interface).
Cadd_class_gen <- nCompile_nClass(add_class, interface = "generic")
obj <- Cadd_class_gen()
# To access fields and methods in an object from the generic interface, 
# use value() and method().
value(obj, 'Cnumber') <- 5
value(obj, 'Cnumber')
method(obj, 'add_Cnumber')(2:4)
```

# Other capabilities

See other User Manual pages to learn about:

1. Generating C++ into package source code.
2. Saving and loading C++ objects easily based on *serialization* of fields.
3. What code can be compiled, and how to mix with C++ code.

# Future plans

1. Automatic copying of fields between uncompiled and compiled objects.
2. Support for `nimble`-style automatic determination of fields and additional compilation flexibility.
3. Access to R fields and methods from C++ methods.
