---
title: "nFunctions and nClasses"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

nCompiler provides new kinds of functions and classes that can be fully or partially compiled.  They can also be used uncompiled (i.e. in R), which is useful for debugging.

# `nFunction`

An `nFunction` is created from an R function and information about argument types and the return type.

```{r}
library(nCompiler)
```

```{r}
add2p1 <- function(x) {
  ans <- x+2.1
  return(ans)
}

nf_add2p1 <- nFunction(
  fun = add2p1,
  argTypes = list(x = 'numericVector()'),
  returnType = 'numericVector()'
)
nf_add2p1(1:3)
```

This can be compiled by:
```{r}
# One may first see ouput from building a local library called nCompLocal.
Cnf_add2p1 <- nCompile_nFunction(nf_add2p1)
Cnf_add2p1(1:3) # The compiled version works.
```

The vector addition is accomplished using the Eigen library in the code-generated C++.

An `nFunction` can also be created by writing type information in the format of default arguments and providing an explicit `returnType` statement anywhere in the function body.  For example:

```{r}
nf_add2p1 <- nFunction(
  fun = function(x = numericVector()) {
    returnType(numericVector())
    ans <- x + 2.1
    return(ans)
  }
)
```

# `nClass`

`nClass`es are similar to reference classes and R6 classes, but they can have two categories of fields and methods: those to be handled via C++ compilation and those that will remain pure R.  A `nClass` is implemented as an R6 class and so follows it most closely.

```{r}
add_class <- nClass(
  classname = "add_class",
  Rpublic = list( ## only available in R
     Rnumber = NULL,
     add_Rnumber = function(v) {
       v + Rnumber
     }
     # initialize not supported yet b/c it needs special handling for full interface.
  ),
  Cpublic = list( ## available in R or compiled version
    Cnumber = 'numericScalar()', #member data -- must have a type declaration
    # member function -- written as an nFunction
    add_Cnumber = nFunction(
      fun = function(v = numericVector()) {
        returnType(double(1))
        ans <- v + Cnumber
        return(ans)
      })
  )
)
# This can be used uncompiled.
```

This can be used uncompiled for debugging purposes.  Then it can be compiled.  For objects of the compiled class, there are two kinds of interfaces, generic and full.  In a generic interface, there isn't an R6 object for every compiled object.  This saves a lot of R processing and memory in large collections of compiled objects.  In a full interface, there is an R6 object for every compiled object.  The full interface is better for nice use of a small number of objects.
```{r}
Cadd_class <- nCompile_nClass(add_class)
# By default we get a full interface, which
# behaves as one would expect for a class generator.
obj <- Cadd_class$new()
obj$Cnumber <- 3  ## This automatically stores into the C++ object
obj$Cnumber
obj$add_Cnumber(2:4) ## and automatically calls the C++ method.
# We can also make a generic interface (or get it out of the full interface).
Cadd_class_gen <- nCompile_nClass(add_class, interface = "generic")
obj <- Cadd_class_gen()
# To access fields and methods in an object from the generic interface, 
# use value() and method().
value(obj, 'Cnumber') <- 5
value(obj, 'Cnumber')
method(obj, 'add_Cnumber')(2:4)
```

# Other capabilities

See other User Manual pages to learn about:

1. Generating C++ into package source code.
2. Saving and loading C++ objects easily based on *serialization* of fields.
3. What code can be compiled, and how to mix with C++ code.

# Future plans

1. Automatic copying of fields between uncompiled and compiled objects.
2. Support for `nimble`-style automatic determination of fields and additional compilation flexibility.
3. Access to R fields and methods from C++ methods.
