---
title: "Parallelization"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

nCompiler uses Threading Building Blocks (TBB), a parallel computing C++ library from Intel.  This is the same library adopted by `RcppParallel`.  Similarly to the way nCompiler uses the Eigen, CppAD, and cereal packages, nCompiler uses TBB by generating C++ for it from simple `nFunction` code.

TBB manages thread creation, saving the programmer a lot of hassle.

Everything in TBB is class (functor) based.  nCompiler generates appropriate classes from `nClass` definitions.

This system has a working prototype.  More complete support for nCompiler types and TBB tools remains for future work.

## Example

Here is a working example.  The name `parallel_for` matches the name of the TBB function.  In the `nFunction` method of the `nClass` below, the first three arguments to `parallel_for` are like `for(i in 1:10) {y[i] <- 2 * x[i]}`.  The next two arguments gives variable(s) that should be copied for each thread (`x`) and variables that can be shared across threads (`y`).  In this example, `x` does not really need to be copied for each thread, but I wanted to put some variable in this role for the example.

The name and syntax of `parallel_for` are up for discussion and may change.

```{r}
library(nCompiler)
nc <- nClass(
  Cpublic = list(
    go = nFunction(
      fun = function(x = 'numericVector') {
        y <- x
        parallel_for(i, 1:10,
                     {y[i] <- 2 * x[i]},
                     "x", ## copy for each thread
                     "y") ## share across threads
        return(y)
      },
      returnType = 'numericVector'
    )))

Cnc <- nCompile(nc)
Cnc1 <- Cnc$new()
## This runs in parallel:
Cnc1$go(101:110)
```

# Future plans

- Support more TBB idioms, such as `parallel_reduce` (drafted).
- Allow more fine-grained creation of TBB functor classes for advanced users.
- Ensure that all types supported by nCompiler work as either copied or shared variables.
- Support fine-grained control of TBB's thread management (perhaps via RcppParallel).
