---
title: "Saving and Loading"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
```

Saving and loading C++ objects used from R is a challenge because those objects are not handled by R's simple `save` function.  `nCompiler` uses the deliciously named `cereal` C++ package to accomplish serialization to support saving and loading C++ objects. 

Serialization refers to stringing the contents of an object together in some standard way for a purpose such as writing it to disk and later reading it.  R's `save` and `load` use serialization of R objects.

`nClass` definitions can have `cereal` code automatically included.  The system is not fully developed,  but the core idea of code-generating `cereal` code into `nClass` definitions works.

Saving and loading compiled objects needs to go hand in hand with re-using generated-and-compiled C++ code.  The latter is done by making use of R's packaging system.  See ["Using nCompiler code in packages"](use_in_packages.html).  With both of these features, one can load a package with the compiled `nClass` definitions and then load saved objects of those classes.

### Example:

Here is an example.  Note that the system is sufficiently incomplete that it must be turned "on" via an option.

```{r}
library(nCompiler)
nc1 <- nClass(
  Cpublic = list(
    Cv = 'numericScalar',
    Cx = 'integerScalar',
    Cfoo = nFunction(
      fun = function(x) {
        return(x+1)
      },
      argTypes = list(x = 'numericScalar'),
      returnType = 'numericScalar')
  )
)
set_nOption('serialize', TRUE)
nc1_generator <- nCompile_nClass(nc1, interface = "generic")
nc1_obj <- nc1_generator[[1]]()
method(nc1_obj, "Cfoo")(1.2) # The method works.
value(nc1_obj, "Cv") <- 1.23 # Access to fields works.
value(nc1_obj, "Cv") 
serialized <- serialize_nComp_object(nc1_obj) #serialize! 
# look at the gooey internals to see that they are a binary
## representation of nc1_obj.
serialized$serialized
## One could save and load the serialized content and then
## reconstitute the object by deserialization:
nc1_restored_obj <- deserialize_nComp_object(serialized) #deserialize!
value(nc1_restored_obj, "Cv") # The value was restored.
method(nc1_restored_obj, "Cfoo")(1.2) # The method works
```

### Future plans:

- support for use of `cereal` with more kinds of C++ member data.
- create some coordination of R's serialization of environments that hold external pointers to nCompiler's C++ objects and `cereal` serialization of the C++ objects. 
- create reasonable behavior for saving and loading from either full or generic interfaces.
- manage serialization/deserialization from multiple DLLs.
