---
title: "Types and argument passing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# Types

Types in `nCompiler` must be static.  These include `nClass` fields and  `nFunction` variables that will be compiled.  (A compileable `nClass` method is an `nFunction`.)  To accomplish static typing, the programmer must declare types of arguments, return objects, and compiled fields (in an `nClass`).  The same variable cannot be used for more than one type.

## Numeric, integer and logical types

These represent three core types of scalar and hence elements of non-scalars (vectors, matrices, arrays).  

Note that in R, "numeric" sometimes means "double *or* integer" [e.g., `is.numeric`] and sometimes means "double" [e.g. `as.numeric`].  In `nCompiler`, "numeric" means "double".

There are several ways to declare these three core types.  We use numeric as an example:
```{r eval = FALSE}
numericScalar()
numericVector()
numericMatrix()
numericArray(nDim = 3)
## nCompiler will also provides some alternatives and ways to
## control sizes and values in type declarations.
```

In these formats, the "`numeric`" prefix can be replaced with "`integer`" or "`logical`".

## How to use numeric, integer, and logical type declarations

Say we want to do compiled vector addition:
```{r}
addVectors <- nFunction(
  fun <- function(x, y) {
    ans <- x + y
    return(ans)
  },
  argTypes = list(x = 'numericVector()', 
                  y = 'numericVector()'),
  returnType = 'numericVector()'
)
# We could compile like this:
# test <- nCompile_nFunction(addVectors)
```

As an alternative, types can be declared directly in code.  The following is an equivalent way to write `addVectors`:
```{r}
addVectors <- nFunction(
  fun <- function(x = numericVector(),
                  y = numericVector()) {
    ans <- x + y
    return(ans)
    returnType(numericVector())
  })
# We could compile like this:
# test <- nCompile_nFunction(addVectors)
```

## Future plans

- Allow type declarations to establish default argument values.
- Allow type declarations to evaluate to new variables in a function body (similar to `nimble`).
- Additional fine-grained control over initialization of values (similar to `nimble`).
- Support access to R objects using Rcpp.
- Support pass-by-copy and pass-by-reference semantics.
- Support sparse matrices.

# Argument passing: pass-by-copy and pass-by-reference

In `nimble`, arguments are passed by value in uncompiled execution and by reference in compiled execution.  This is internally inconsistent, and the compiled behavior is efficient but not R-like.

In `nCompiler`, we aim to support consistent type behavior and let the user declare what they need.  Three options are:

1. Pass by value.
2. Pass by reference, with the entire object mutable, including size.  This would work for passing an entire object such as a matrix `X`.  `X` could be resized in the called function.
3. Pass by reference, with values but not sizes mutable.  This would work for passing an indexed subset of an object, like `X[2:5, 3:6]`.

Note that pass-by-reference is desirable for efficient C++, and after all a goal of C++ is efficiency.  However, it is very un-R-like.  To achieve consistent uncompiled and compiled behavior for debugging purposes, we can impose non-standard evaluation in R via modification of the function body.

Currently, a system for passing entire objects by reference in R has been implemented.  There are two ways to declare pass-by-reference: the `refArgs` argument, and wrapping the type declaration in `ref()`.

```{r eval=TRUE}
library(nCompiler)
foo <- nFunction(
  fun = function(a = numericScalar()) {
    a <- a + 1
    a
  },
  refArgs = 'a', # Make a pass-by-reference
  returnType = quote(numericScalar())
)
a <- 10
foo(a)
a # a was modified by foo, becuase it was passed by reference
```

Achieving pass-by-reference semantics across the R-C++ boundary will be an issue.  For example, when calling a compiled function from R rather than from another compiled function.

## Future plans

- Implement these approachs in R and code-generated C++.
