---
title: "Types and argument passing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```

# Types

Types in `nCompiler` must be static.  These include `nClass` fields and  `nFunction` variables.  To accomplish static typing, the programmer must declare types of arguments, return objects, and Cpublic (compilable) fields in an `nClass`.  The same variable cannot be used for more than one type.

## Numeric, integer and logical types

These represent three core types of scalar and hence elements of non-scalars (vectors, matrices, arrays).  

Note that in R, "numeric" sometimes means "double *or* integer" [e.g., `is.numeric`] and sometimes means "double" [e.g. `as.numeric`].  In `nCompiler`, "numeric" means "double".

There are several ways to declare these three core types.  We use numeric as an example:
```{r eval = FALSE}
numericScalar()
numericVector()
numericMatrix()
numericArray(nDim = 3)
## nCompiler will also provides some alternatives and ways to
## control sizes and values in type declarations.
```

In these formats, the "`numeric`" prefix can be replaced with "`integer`" or "`logical`".

## How to use numeric, integer, and logical type declarations

Say we want to do compiled vector addition:
```{r}
library(nCompiler)
addVectors <- nFunction(
  fun <- function(x, y) {
    ans <- x + y
    return(ans)
  },
  argTypes = list(x = 'numericVector()', 
                  y = 'numericVector()'),
  returnType = 'numericVector()'
)
# We could compile like this:
# test <- nCompile(addVectors)
```

When using character-string type declarations, the `()` is optional if no arguments are needed.  Simply `'numericVector'` would work above.

As an alternative, types can be declared directly in code.  The following is an equivalent way to write `addVectors`:
```{r}
addVectors <- nFunction(
  fun <- function(x = numericVector(),
                  y = numericVector()) {
    ans <- x + y
    return(ans)
    returnType(numericVector())
  })
```

nCompiler also provides backward-compatibility support for nimble-style type declarations such as `double(1)` for a numeric vector.

## Future plans

- Allow type declarations to establish default argument values.
- Allow type declaration code to work as a regular function (similar to `nimble`).
- Additional fine-grained control over initialization of values (similar to `nimble`).
- Support access to R objects using Rcpp.
- Finish support for pass-by-copy and pass-by-reference semantics.
- Support sparse matrices.

# Argument passing: pass-by-copy and pass-by-reference

In `nimble`, arguments are passed by value in uncompiled execution and by reference in compiled execution.  This is internally inconsistent, and the compiled behavior is efficient but not R-like.

In `nCompiler`, we aim to support consistent type behavior and let the user declare what they need.  Three options are:

1. Pass by value.
2. Pass by reference, with the entire object mutable, including size.  This would work for passing an entire object such as a matrix `X`.  `X` could be resized in the called function.
3. Pass by block reference, which means values but not sizes are mutable.  This would work for passing an indexed subset (block) of an object, like `X[2:5, 3:6]`.

Note that pass-by-reference is desirable for efficient C++, and after all a goal of C++ is efficiency.  However, it is very un-R-like.  To achieve consistent uncompiled and compiled behavior for debugging purposes, we programmatically modify the R function body.

Currently, a system for passing entire objects by reference in both uncompiled (R) and compiled (C++) code has been implemented.  (It remains to implement pass-by-reference semantics when calling C++ from R.)

There are two ways to declare pass-by-reference: the `refArgs` argument, and wrapping the type declaration in `ref()`.

## Pass-by-reference in uncompiled code

Note that this represents un-R-like behavior and has been implemented to support matching uncompiled and compiled behavior.

```{r eval=TRUE}
library(nCompiler)
f1 <- nFunction(
  fun = function(x = 'numericVector',
                 y = 'numericVector') {
    x <- x + 1
    ans <- x + y
    return(ans)
  },
  refArgs = 'x',
  returnType = 'numericVector'
)
x <- 1:3
f1(x, 11:13)
x # x was modified by f1, because it was passed by reference
```

## Pass-by-reference in compiled code

To see pass-by-reference work in compiled code, we need a second function to call the first.

```{r}
f2 <- nFunction(
  fun = function(x = 'numericVector',
                 y = 'numericVector') {
    f1(x, y)  # We are not interested in the returned value.
    return(x) # We are interested in seeing that x was modified.
  },
  returnType = 'numericVector'
)
compiled <- nCompile(f2, f1)
x <- 1:3
compiled$f2(x, 11:3) # We see the modified x returned.
```

## Future plans

- Finish support for pass-by-reference.
- Add support for passing by block references.
