---
title: "Use in packages"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Using nCompiler code in packages

nCompiler has partially-built support for generating code into packages.  This includes:

- generating C++ code into src directory and using Rcpp to compile it into the package.
- generating C++ code into inst directory so that it can be included from other packages.
- generating R code into R directory with `nClass` definitions to interface with compiled `nClass` objects.

Note that these features work for simple cases but are not fully developed yet.

Following is an example of how to do so for an `nClass`.
```{r}
library(nCompiler)
nc1 <- nClass(
  classname = "nc1",
  Rpublic = list(
    p1 = function(x) x+1
  ),
  Cpublic = list(
    x = 'numericScalar',
    cp1 = nFunction(
      fun = function(x = numericScalar()) {
        ans <- x+1
        return(ans)
        returnType(numericScalar())
      }
    )
  )
)
package_dir <- tempdir() # Normally one would choose a local directory
# A variety of output from pkgKitten and Rcpp is generated
writePackage(
  nc1,
  dir = package_dir,
  package.name = "nc1Package",
  clean = TRUE
)
buildPackage("nc1Package",
             dir = package_dir)
```

## Load and use nc1Package

```{r}
library("nc1Package")
my_nc1 <- nc1Package::nc1$new()
my_nc1$cp1(1)
```



# Future plans

- Support multiple objects, exporting, finding needed libraries during package building, and other basic package controls.
