---
title: "Use in packages"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Using nCompiler code in packages

nCompiler has support for generating code into packages.  This includes:

- generating C++ code into src directory and using Rcpp to compile it into the package.
- generating C++ code into inst directory so that it can be included from other packages.
- generating R code into R directory with `nClass` definitions to interface with compiled `nClass` objects.
- generating documentation for compiled features into the appropriate R and C++ files and controlling exporting.

There are three main functions dealing with package writing in nCompiler:
`writePackage`, `buildPackage`, and `erasePackage.`

`writePackage` handles the creation of source code. It takes one or more nFunction
and nClass generator objects and some package information and writes the relevant
code to the specified directory. This function can also add to or modify an
existing package.

`buildPackage` handles package building and installation, including compilation
of generated C++ code.

`erasePackage` wipes and deletes a specified package when it's no longer needed.

Following is an example of how to do so for an `nClass`.
```{r}
library(nCompiler)

nc1 <- nClass(
  classname = "nc1",
  Rpublic = list(
    p1 = function(x) x+1
  ),
  Cpublic = list(
    x = 'numericScalar',
    cp1 = nFunction(
      fun = function(x = numericScalar()) {
        ans <- x+1
        return(ans)
        returnType(numericScalar())
      }
    )
  )
)
package_dir <- tempdir() # Normally one would choose a local directory
# A variety of output from pkgKitten and Rcpp is generated
nWritePackage(
  nc1,
  dir = package_dir,
  package.name = "nc1Package"
)
nBuildPackage("nc1Package",
             dir = package_dir)
```

## Load and use nc1Package

```{r}
library("nc1Package")
my_nc1 <- nc1Package::nc1$new()
my_nc1$cp1(1)
```
The source code for nc1package could be erased, and the library uninstalled,
with `erasePackage`.

```{r}
if (FALSE) nErasePackage(dir = package_dir, package.name = "nc1Package")
```

## Documenting features

We can add a new element with accompanying documentation.

```{r}
foo <- nFunction(
  name = "foo",
  fun = function(x = numericScalar()) {
    ans <- x+2
    return(ans)
    returnType(numericScalar())
  }
)
rox <- nDocumentNFunction(
        name  = "foo",
        title = "A Test nFunction",
        description = "This nFunction just adds 1 to a 
                       scalar input.",
        params = list(x = "A scalar to which 1 will be added."), 
        otherRoxygen = "//' @export", 
        checkAgainstObj = FALSE)

nWritePackage(foo, 
             package.name = "nc1Package", 
             dir = package_dir,
             roxygen = list(foo = rox),
             modify = TRUE) # modify = TRUE lets us add content

nBuildPackage("nc1Package",
             dir = package_dir)

```

To see the output of the modified nc1Package, we would need to start a new session.

# Future plans

- Finding needed libraries during package building, and other basic package controls.
