---
title: "types for inputs and outputs"
format: html
editor: visual
---

## Quick summary

- You need to say what the types of inputs and outputs will be.
- Usually the types of variables created within functions will be handled automatically.
- Types include basic types (scalars, vectors, matrices, or arrays of doubles, integers, or logicals), `nClass` types, `nList` types, and `Rcpp` types, among others.

::: {.callout-important collapse="true"}
## nCompiler supports true scalars

In R, there are no true scalar. Rather, what feel like scalars are length-1 vectors. In `nCompiler`, generated C++ can include true scalars, so there is an important type distinction between true scalars and non-scalars.
:::


## How to use type declarations

Type declarations are needed to cross the threshold between R and C++. R allows variables to be any type (dynamic typing), while C++ requires each variable to have a declared, unchanging type (static typing).


### Two places for type declarations

Type declarations can be given:

- directly in the function code, or
- in separate arguments to `nFunction`. 

The `nf_mult2` example given above used the second approach. Here is the alternative:

```{r eval=FALSE}
nf_mult2 <- nFunction(
  fun = function(x = 'numericVector') {
    ans <- 2 * x
    return(ans)
    returnType('numericVector') # This can appear anywhere.
  })
```

It is also fine to have text like a function call, e.g. `'numericVector()'` because sometimes one wants to include arguments.

The `returnType` call can appear anywhere in the function body. It does not affect code execution.

### Two syntaxes for types

Types declarations can be given as:

- character strings, or
- code (sometimes "quoted").

The examples so far use character strings. Here are examples with code:

```{r eval=FALSE}
nf_mult2 <- nFunction(
  fun = function(x = numericVector()) {
    ans <- 2 * x
    return(ans)
    returnType(numericVector())
  })
```

or

```{r eval=FALSE}
nf_mult2 <- nFunction(
  fun = mult2,
  argTypes = list(x = quote(numericVector())),
  returnType = quote(numericVector())
)
```

### Type objects

Sometimes it is useful to make an object that holds the type, allowing one to write code to construct `nFunctions`. That can be done with the function `nMakeType` and the special syntax `T` in type declarations. For example:

```{r eval=FALSE}
my_type <- nMakeType(numericVector())
nf_mult2 <- nFunction(
  fun = function(x = T(my_type)) {
    ans <- 2 * x
    return(ans)
  },
  returnType = quote(T(my_type))
)
```

For illustration, this uses `my_type` once in the function code and again in the `returnType` argument.

::: {.callout-caution collapse="true"}
## The system for type objects may change.

The scheme for making and using objects containing type information is still being designed and so might change.
:::

### Numeric types

A numeric type in `nCompiler` comprises the number of dimensions and the scalar type of each element. A special type for sparse matrices is also provided.

-   "Number of dimensions" really means the number of *index* dimensions. For example a matrix or 2D array has two (index) dimensions and a 3D array has three index dimensions. [^types-1]
-   The scalar type of each element can be `numeric`, `integer`, or `logical`. A synonym for `numeric` is `double` (for standard "double precision" numbers). Note that in R, "numeric" means "integer or double" (e.g. try `is.numeric`), but in `nCompiler` type declarations, `numeric` means `double`.

[^types-1]: Of course, mathematically a vector of length n represents a point in n-dimensional space. That's not what we mean by number of dimensions. A vector is considered one-dimensional because elements are identified by just one index.

#### Declaring numeric types

There are several ways to declare numeric types:

- `integerScalar()`, `numericScalar()`, and `logicalScalar()` are what they sound like.
- `nScalar(type="integer")` is the same as `integerScalar()`. The `type` argument defaults to `"double"` and can also be `"logical"`. (The "n" in "`nScalar`" and similar names below stands for "nCompiler", not "numeric".)
- `numericVector()`, `nVector()`, and `double(1)` are all double-precision vectors.
- `integerVector()`, `nVector(type="integer")` and `integer(1)` are all integer vectors.
- `logicalVector()`, `nVector(type="logical")` and `logical(1)` are all logical vectors.
- The `*Vector()` forms and `nVector` can take a `length` argument. `nVector` can also take a `value` argument. `double(1)` exists for compatibility with `nimble`'s type system. The forms like `double(1)` exist for compatibility with `nimble`'s type system.
- The prefixes "`numeric`", "`integer`", and "`logical`" can also go with the suffixes "`Matrix`" or "`Array`". For example: `integerArray(nDim=3)`.
- `nMatrix` and `nArray` are similar but take the element type (`double`, `integer`, or `logical` are an argument).

### Character types

- `string`

### Sparse matrix types

- `nSparseMatrix`
- `nSparseVector`

### Rcpp types

Many Rcpp types are supported. Using these requires some of your own C++ coding. These include:

- `RcppEnvironment`
- `RcppList`
- `Rcpp<Numeric, Integer, Character, Logical, Complex><Vector, Matrix>` (e.g. `RcppCharacterVector`).
- Others to be listed later.

### C++ types

- `SEXP`
- `void`
- `cppVar`
- `nCpp` (for direct coding of an arbitrary type)

### nClass and nList types

These will be covered in later sections.
